---
title: "Assignment B-1"
output: md_document
---

STAT 545 Assignment B1
=========
Brian Deng
60169810

## Loading libraries & data
```{r}
library(tidyverse)
library(roxygen2)
library(testthat)

library(datateachr)
library(palmerpenguins)
library(gapminder)
```


# Exercise 1 & 2: Making and Documenting the Function

During my mini data analysis project, I coded quite a few histograms. 
I want to create a function that takes in data inputs and outputs a histogram.
This way, I don't need to type out the histogram code each time I want a histogram. 
I like magenta so the function will plot a histogram with magenta coloured bars. 

```{r}
#' Plot a histogram with magenta bars of a numeric variable
#' 
#' This function plots a simple histogram using ggplot2 with magenta coloured bars. 
#' The user inputs the data frame, column name, bin number.
#'
#' @param data The data frame that contains the variable of interest. Mandatory.
#' @param variable The variable of interest, which must be numerical. Mandatory. 
#' @param bins The number of bins for the histogram, which must be a positive integer. Mandatory. 
#'
#' @return A histogram plot with the variable of interest as the x axis. 
#' @examples
#' plot_magenta_histogram(cancer_sample, concavity_mean, 50)
#' plot_magenta_histogram(cancer_sample, radius_mean, 100)
#' plot_magenta_histogram(penguins, bill_length_mm, 50)
#' plot_magenta_histogram(gapminder, lifeExp, 100)
#'
#' @export

plot_magenta_histogram <- function(data, variable, bins) {

  # check data frame input
  if (!is.data.frame(data)) {
    stop("The input data frame must be a data frame.")
  }
  
  # dropping rows that contain NA's
  data <- filter(data, !is.na({{ variable }}))
  
  # check numeric variable input after filtering
  if (nrow(data) == 0) {
    stop("The input variable does not contain data after filtering")
  }

  # check numeric variable input
  if (!is.numeric(pull(data, {{ variable }}))) {
    stop("The input variable must be a numeric variable.")
  }
  
  # check bin input
  if (!is.numeric(bins) | bins <= 0 ) {
    stop("The input bins must be a positive integer")
  }
  
  # dropping rows that contain NA's
  data <- filter(data, !is.na({{ variable }}))
  
  # plotting histogram
  ggplot(data, aes(x = {{ variable }})) +
      geom_histogram(bins = bins, fill = "magenta", colour = "pink") +
      theme_bw()
}
```


# Exercise 3: Examples
I'm now going to provide some examples using various datasets. 

Plotting the distribution of average radius of tumour samples
```{r}
# the cancer_sample dataset has columns of numerical variables 
plot_magenta_histogram(cancer_sample, radius_mean, 50)
```

Example of common errors with this code

```{r, error = TRUE}
# mistyping the variable name will simply return an error
plot_magenta_histogram(cancer_sample, raduis_maen)

# forgetting to specify the number of bins
plot_magenta_histogram(cancer_sample, concavity_mean)

# putting a negative integer as the number of bins
plot_magenta_histogram(cancer_sample, concavity_mean, -1)

# putting a non-integer as the number of bins 
plot_magenta_histogram(cancer_sample, concavity_mean, "50")

# inputting a non-data frame 
plot_magenta_histogram(cancer_sample$diagnosis, concavity_mean)

# inputting a non-numerical variable 
plot_magenta_histogram(cancer_sample, diagnosis)
```

Examples of the code with another dataset

```{r}
# plotting the penguin bill length distribution in all penguins
plot_magenta_histogram(gapminder, gdpPercap, 100)

# plotting the life expectancy from countries around the world
plot_magenta_histogram(gapminder, lifeExp, 100)
```

Example of variable that contains a lot of NA's

```{r}
# all columns some NA's in the column 
glimpse(penguins)

# in all the plots below, the NA's are dropped and do not show up in the histogram
plot_magenta_histogram(penguins, bill_length_mm, 50)
plot_magenta_histogram(penguins, bill_depth_mm, 50)
plot_magenta_histogram(penguins, flipper_length_mm, 50)
plot_magenta_histogram(penguins, body_mass_g, 50)
```

# Exercise 4: Testing the Function

Before we start, I'm going to create data frames to run tests from
```{r}
# a data frame with multiple columns with various types of variables 
test_data <- data.frame(
  diagnosis = c("M", "M", "M", "B", "B", "B"), 
  radius = c(100, 101, 95, 34, 56, 21),
  radius_non_numerical = c("100", "101", "95", "34", "56", "21"),
  radius_NA = c(100, 101, 95, 34, NA, 21)
)

# a data frame with no data
test_data_zero <- data.frame(
  zero = numeric(0)
)

# a data frame with NA data
test_data_NA <- data.frame(
  NAs = c(NA, NA, NA, NA)
)
```


Test that the function will not take in a non-numeric variable

```{r}
# the diagnosis column contains non-numerical values
test_that("Function will error because a non-numeric variable was inputted", {
  expect_error(plot_magenta_histogram(test_data, diagnosis, 100))
})

# the radius_non_numerical column also contains non-numerical values
test_that("Function will error because a non-numeric variable was inputted", {
  expect_error(plot_magenta_histogram(test_data, radius_non_numerical, 100))
})
```

Test that the function will not error if there are NA's present 

```{r}
# the NA values will be filtered out 
test_that("Function will be able to handle NA values", {
  expect_no_error(plot_magenta_histogram(test_data, radius_NA, 100))
})
```

Test that the function errors if there is nothing in the data frame

```{r}
# the data frame has no data in it
test_that("Function will error when there is nothing in the data frame", {
  expect_error(plot_magenta_histogram(test_data_zero, zero, 100))
})

# since the data frame only has NA values, they will be filtered out and the data frame will no longer have data 
test_that("Function will error when there is nothing in the data frame", {
  expect_error(plot_magenta_histogram(test_data_NA, NAs, 100))
})
```

Test that a proper ggplot object is outputted

```{r}
test_that("ggplot structure is correct", {
  plot <- plot_magenta_histogram(penguins, bill_length_mm, 50)
  
  # check that the plot is a ggplot object
  expect_s3_class(plot, "ggplot")
  
  # check that the input variable is correctly mapped
  expect_equal(rlang::as_name(plot$mapping$x), "bill_length_mm")
 })
```

